Molecular dynamics simulation of synchronization in driven particles

### 2021-01-12 manuscript/manuscript_ajp_2020.tex
write about new fig. 3 make exercise to explore backwards motion

A single particle %on this landscape
responds to the applied driving force
by synchronizing its motion across the landscape %maxima 
to the periodic pattern of 
applied driving force.
%along the y-direction
%({\it redundant, just refer back to eq. in Model?})
%and $V_{0y}=0.2$.
%This is illustrated in Fig.~\ref{fig:1}(a) where
%the red (blue) regions show local maxima (minima).
%
The numerical implementation of the landscape 
is calculated with Eq.~\ref{eq:ysubstrate} as 
\begin{equation}
  \label{eq:force}
  F^l_y(y) = -A_{p} \sin{(N_p \pi y / L)} 
\end{equation}
where the force is scaled with parameter $A_{p}$.
In this section, we fix the landscape parameters
to $A_{p} = 0.1$ force units
with $N_p=20$ minima in the landscape
corresponding to a spatial period $\lambda = 1.825 a_0$.

In Fig.~\ref{fig:0} 
a single driven particle adjusts
its motion so the particle traverses
the periodic landscape with the same rhythm as 
the applied time-dependent force $F^d(t)$.
%
In Fig.~\ref{fig:0}(a)
we plot $F^d(t)$ defined in Eq.~\ref{eq:drive}.
%as a function of time.
The time independent constants are 
$F^{ac}=0.05$ and $f=0.01$ cycles per time unit.
The temporal period of the driving force is
$T = 1/f = 100$ time units.
To avoid transient oscillations at early times, 
the dc driving force is initially zero   
and 
slowly increased %$F^{dc}$.
at a rate of $\Delta F^{dc} = 0.01$ every $\Delta t = 4$ time units  
to the constant value $F^{dc}=0.1$.
In Fig.~\ref{fig:0}(a) we plot $F^{dc}$
in blue to show the increase in 
$F^{dc}$ for the first $40$ time units
In orange we show 
the combined ac and dc drive $F^d(t)$ 
when the magnitude of the ac applied force is initially zero.
%since $\sin{(0)}=0$.
We mark the maxima of $F^{ac}(t)$
with the
vertical dashed lines %in panels (a) and (b)
(i.e. $\sin{(\omega t)} = 1$).

@@

In Fig.~\ref{fig:0}(b) 
we show the $y-$position of the particle
as a function of time,
where we 
normalize $y$ by $\lambda$, the spatial period of the potential.
The initial particle position is $y = 1$,
where the landscape force is negative $F^l < 0$,
causing the particle to move into
a landscape minima at the start of the simulation.
As the driving force increases,
%the effects of the landscape become less obvious,
and the motion of the particle
becomes locked to the driving period $T$.
The horizontal dashed lines
coincide with the condition the substrate force is
minimum $F^l = -A_p$ 
(i.e. $\sin{(\pi y / \lambda)} = 1$).
%or $\pi y / \lambda = m \pi/2$, where $m = 1, 5, 7$, {\it etc.}
Once the particle motion
coincides with the applied drive, 
the particle position $y$
crosses the intersection of
the horizontal and vertical lines.
This indicates 
the driving force is maximum positive when the
force on the particle from the landscape is maximum negative,
indicating synchronization.
The slope of $y/\lambda$
is proportional to the net force on the particle.
This can be seen in 
Fig.~\ref{fig:0} %we demonstrate this relationship,
where local extrema appear in $y/\lambda$
when $F_{D}(t) \approx 0 $. 
We invite interested students to repeat this
simulation for any initial particle position %of $y$
and a range of simulation constants
to demonstrate that synchronization in this model
does not depend on a particular set of conditions.
%but is a direct consequence of the simplicity of the model.

@@

The competition between the driving force and landscape potential
can produce a variety of hopping patterns in the particle motion.
If $A_p$ is large compared to the extrema of $F^{d}(t)$,
the particle is trapped in a single
minima so it oscillates back and forth with no net velocity.
When $F^d(t) > A_p$, a particle can 
overcome the barrier height of the landscape maxima,
and 
the particle hops between minima in the energy landscape.
In Fig.~\ref{fig:0}
the particle moves
in the positive y-direction
through $\Delta y = 2 \lambda$ in one time period $\Delta t = T$,
with 
the average velocity 
$\bar{v}_y = 2 \lambda f$ where $T = 1/f$.
%reaching a minimum position
%of $y/\lambda = 1$
%over the total period $T = 100$.
%The average velocity $\bar{v}_y$
%is the average displacement $\Delta y = y(t_0+T) - y(t_0)$
%over the period of the driving force.
%In Fig. ~\ref{fig:0}
%the average displacement during one time period is a single wavelength
%of the substrate $y(t_0+T) - y(t_0) = \lambda$.
%single dc case with figure
%dc_current 0.01
%ac_current 0.05
%ac_frequency 0.01
%fp_y 0.1

The ratio of $F^{ac}$, $F^{dc}$ and the barrier height $A_p$
control the rate and length of the particle hops
forward and backward in the landscape.
When $F^{ac}/F^{dc} > 1$
and $F^{dc} - F^{ac} > A_p$ the particle
moves in the negative y-direction 
during part of the cycle.
To explore the possible hopping patterns,
we sweep through a range of $F^{dc}$ for fixed $F^{ac}$ and $A_p$,
as shown in Fig.~\ref{fig:1}(b).
We keep $F^{dc}$ constant
for $\Delta t = 10^4$ timesteps
and 
sample the average velocity $\langle v_y \rangle $ 
as a function of $F^{dc}$.

@@

Our simulations reproduce results presented in 
Juniper {\it et al.} \cite{Juniper2015, Juniper2018}
which demonstrated
mode locking in
experiments of 
driven colloids on a
optical periodic landscape.

@@

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{single}
\caption{(Average particle velocity in the y-direction $\langle v_{y} \rangle$ as a function of a constant driving force $F^{dc}$.}
\label{fig:1}
\end{figure}
\end{center}

>>>>

A single particle %on this landscape
responds to the applied driving force
by moving across the landscape %maxima
synchronized with 
the period of $F^d(t)$.
%applied driving force.
%along the y-direction
%({\it redundant, just refer back to eq. in Model?})
%and $V_{0y}=0.2$.
%This is illustrated in Fig.~\ref{fig:1}(a) where
%the red (blue) regions show local maxima (minima).
%
The numerical implementation of the landscape 
is calculated with Eq.~\ref{eq:ysubstrate} as 
\begin{equation}
  \label{eq:force}
  F^l_y(y) = -A_{p} \sin{(N_p \pi y / L)} 
\end{equation}
where the force is scaled with parameter $A_{p}$.
When $F^d(t) > A_p$, a particle can 
overcome the barrier height of the landscape,
and 
the particle hops between minima in the energy landscape.
In this section, we fix the landscape parameters
to $A_{p} = 0.1$ force units
with $N_p=20$ minima in the landscape
corresponding to a spatial period $\lambda = 1.825 a_0$.

In Fig.~\ref{fig:0} 
the driven particle adjusts
its motion to traverse
the periodic landscape with the same rhythm as 
the applied time-dependent force $F^d(t)$.
%
In Fig.~\ref{fig:0}(a)
we plot $F^d(t)$ %defined in Eq.~\ref{eq:drive}.
as a function of time with
%The time independent
constants 
$F^{ac}=0.05$ and $f=0.01$ cycles per time unit.
The temporal period of the driving force is
$T = 1/f = 100$ time units.
To avoid transient oscillations at early times, 
the dc driving force is initially zero   
and 
slowly increased %$F^{dc}$.
at a rate of $\Delta F^{dc} = 0.01$ every $\Delta t = 4$ time units  
to the constant value $F^{dc}=0.1$.
In Fig.~\ref{fig:0}(a) we plot $F^{dc}$
in blue to show the increase in 
$F^{dc}$ for the first $40$ time units
In orange we show 
the combined ac and dc drive $F^d(t)$ 
when the magnitude of the ac applied force is initially zero.
%since $\sin{(0)}=0$.
We mark the maxima of $F^{ac}(t)$
with the
vertical dashed lines %in panels (a) and (b)
(i.e. $\sin{(\omega t)} = 1$).

@@

In Fig.~\ref{fig:0}(b) 
we show the $y-$position of the particle
as a function of time.
We 
normalize $y$ by $\lambda$, the spatial period of the potential
to show how the particle moves between substrate minima.  
The initial particle position is $y = 1$,
where the $F^l$ is negative, 
causing the particle to move to
a landscape minima at the start of the simulation.
As the driving force increases,
%the effects of the landscape become less obvious,
the motion of the particle
becomes locked to the driving period $T$.
The particle moves
in the positive y-direction
through $\Delta y = 2 \lambda$ in one time period $\Delta t = T$,
with 
the average velocity 
$\bar{v}_y = 2 \lambda f$. % where $T = 1/f$.
The horizontal dashed lines
coincide with the condition the substrate force is
minimum $F^l = -A_p$ 
(i.e. $\sin{(\pi y / \lambda)} = 1$).
%or $\pi y / \lambda = m \pi/2$, where $m = 1, 5, 7$, {\it etc.}
Once the particle motion
coincides with the applied drive, 
the particle position $y$
crosses the intersection of
the horizontal and vertical lines.
This indicates 
the driving force is maximum positive when the
force on the particle from the landscape is maximum negative,
indicating synchronization.
The slope of $y/\lambda$
is proportional to the net force on the particle.
This can be seen in 
Fig.~\ref{fig:0} %we demonstrate this relationship,
where local extrema appear in $y/\lambda$
when $F_{D}(t) \approx 0 $. 
We invite interested students to repeat this
simulation for any initial particle position %of $y$
and a range of simulation constants
to demonstrate that synchronization in this model
does not depend on a particular set of conditions.
%but is a direct consequence of the simplicity of the model.

@@

The competition between the driving force and landscape potential
can produce a variety of hopping patterns in the particle motion.
If $A_p$ is large compared to the extrema of $F^{d}(t)$,
the particle oscillates back and forth
in a single minima
with no net velocity.  
%In Fig.~\ref{fig:0}

%reaching a minimum position
%of $y/\lambda = 1$
%over the total period $T = 100$.
%The average velocity $\bar{v}_y$
%is the average displacement $\Delta y = y(t_0+T) - y(t_0)$
%over the period of the driving force.
%In Fig. ~\ref{fig:0}
%the average displacement during one time period is a single wavelength
%of the substrate $y(t_0+T) - y(t_0) = \lambda$.
%single dc case with figure
%dc_current 0.01
%ac_current 0.05
%ac_frequency 0.01
%fp_y 0.1

The relative values of $F^{ac}$, $F^{dc}$ and $A_p$
control the rate and length of the particle hops
forward and backward in the landscape.
When $F^{dc} - F^{ac} < -A_p$ the particle
moves in the negative y-direction 
during part of the cycle.
We explore this in Ex.~\ref{ex:backwards}.
To explore the possible hopping patterns,
we sweep through a range of $F^{dc}$ for fixed $F^{ac}$ and $A_p$,
as shown in Fig.~\ref{fig:1}(b).
We keep $F^{dc}$ constant
for $\Delta t = 10^4$ timesteps
and 
sample the average velocity $\langle v_y \rangle $ 
as a function of $F^{dc}$.

@@

Our simulations reproduce results presented in 
Juniper {\it et al.} \cite{Juniper2015, Juniper2018}
which demonstrated
mode locking in
experiments and simulations of 
driven colloids on a
optical periodic landscape.

@@

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{sweep_FDC_vs_vx.pdf}
\caption{(Average particle velocity  $\langle v_{y} \rangle$ as a function of $F^{dc}$.  We fix the amplitude of $F^d(t)$ as $F^{ac}=0.05$ as in Fig.~\ref{fig:0}}.
\label{fig:1}
\end{figure}
\end{center}
